### 2015-12-01 paper/5-related-work.tex
changes to paper, again

%TODO explain about other DEVS simulation kernels, and how they differ
%TODO explain about PythonPDEVS, the most related in terms of feature and "legacy" --> now in C++11 and way better performance
\subsection{PythonPDEVS}
Dxex is closely related to PyPDEVS in design and philosophy. PyPDEVS allows anyone who grasps the DEVS formalisms to immediately simulate his/her model without having to consider the kernel implementation. C++ implementations cannot hope to match the fast prototype/edit/run cycles provided by PyPDEVS, although this can be minimized by building the kernels as libraries. %maybe warn that if a header is touched, you still have to compile from scratch ? (eg. TIME_FP)
Advanced features such as activity based relocation and the performance gains this results in, are still unique to PyPDEVS. \\
\subsection{Adevs}
%TODO explain about adevs, the most related in terms of performance --> we implement more synchronization options and offer better performance because ...
Adevs's source code is still under active development, allowing for an exact comparison in performance and features. It remains in most aspects the fastest simulation engine for the DEVS formalism, but it lacks an optimistic synchronization implementation. %Tracing ?
By virtue of not flattening Coupled Models, performance suffers in increasingly hierarchical models.
%checkme
\subsection{CD++}
%TODO explain about CD++, the most related in terms of synchronization --> we implement everything in a single program, with a non-fragmented code-base
%TODO make mention of Warped, which is the kernel used in CD++, and why we don't use it
Different projects on CD++ offer conservative (CCD++) as well as optimistic (PCD++) parallel simulation. In contrast to our single program, with a non-fragmented code-base, neither projects offer both synchronization protocols. CD++ relies on the WARPED kernel. It is a middleware that provides memory, event, file, time and communication scheduling. WARPED is not used here since we operate explicitly on a shared memory system and since we wanted to design our kernels using the least amount of overhead possible.

>>>>

%TODO explain about other DEVS simulation kernels, and how they differ
%TODO explain about PythonPDEVS, the most related in terms of feature and "legacy" --> now in C++11 and way better performance
\subsection{PythonPDEVS}
Dxex is closely related to PyPDEVS in design and philosophy. PyPDEVS allows anyone who grasps the PDEVS formalisms to immediately simulate his/her model without having to consider the kernel implementation.A Python based implementation offers the advantage of very fast prototype/run/evaluate cycles, this can't be matched by a C++ simulator but it should be noted that once the kernels are compiled (shared libraries), the actual compilation time of the model is small enough to make prototyping possible.
Advanced features such as activity based relocation and the performance gains this results in, are still unique to PyPDEVS.
\subsection{Adevs}
%TODO explain about adevs, the most related in terms of performance --> we implement more synchronization options and offer better performance because ...
Adevs's is still under active development, allowing for an exact comparison in performance and features. It remains one of the fastest simulation engines for the PDEVS formalism, but it lacks an optimistic synchronization implementation. %Tracing ?
By virtue of not flattening Coupled Models, adevs's performance degrades in increasingly hierarchical models. On the other hand, dxex manages lookahead with more overhead than adevs, leading to a performance difference as the model count increases. Finally, adevs employs the full dependency graph, contrasted to dxex kernels which only observe 1-edge removed nodes. 
%checkme
\subsection{CD++}
%TODO explain about CD++, the most related in terms of synchronization --> we implement everything in a single program, with a non-fragmented code-base
%TODO make mention of Warped, which is the kernel used in CD++, and why we don't use it
Different projects on CD++ offer conservative (CCD++) as well as optimistic (PCD++) parallel simulation. In contrast to our single program, with a non-fragmented code-base, neither projects offer both synchronization protocols. CD++ relies on the WARPED kernel. It is a middleware that provides memory, event, file, time and communication scheduling. We did not use the WARPED kernel (nor the underlying MPI), dxex is designed specifically for a shared memory architecture and as such any middleware, while feature-rich, would have lead to unacceptable overhead.

### 2015-12-01 paper/6-conclusions.tex

%TODO conclude the paper, saying that our new simulation tool offers conservative and optimistic synchronization, resulting in good performance, which is even better than adevs
%TODO future work: more things that we might want to implement, and probably something about automatically switching between synchronization protocols at runtime (and reference my activity papers ;))

>>>>

%TODO conclude the paper, saying that our new simulation tool offers conservative and optimistic synchronization, resulting in good performance, which is even better than adevs
Both synchronization implementations offer good performance in differing simulations, in some simulations dxex outperforms adevs whereas in others we can still improve. 
The optimistic implementation uses an aggressive TimeWarp variant, this needs to be extended with lazy evaluation/cancellation to function in cyclic simulations. 
\subsection{Future work}
%TODO future work: more things that we might want to implement, and probably something about automatically switching between synchronization protocols at runtime (and reference my activity papers ;))
\subsubsection{Activity}
As shown in \cite{PythonPDEVS_ACTIMS} activity and allocation of models across kernels is a key aspect in achieving high performance in any parallel implementation. Allocating models so that there are no dependency cycles between their containing kernels is a first step, but not always possible. For optimistic one can use re-allocation to break (runtime dependency) cycles and/or perform load-balancing. If kernels are unevenly balanced they will begin to drift fast, causing increasingly more reverts. There is a limited framework implemented to track model activity for debugging purposes, this could be extended to enable the above strategies.
\subsubsection{Hybrid}
The optimistic implementation could use (null/eot/eit) from conservative to detect and/or reduce the cost of reverts without completely stalling on influencing kernels.
Conservative kernels could be extended with runtime information about influencing models, for example if one can guarantee a static dependency is not used for a fixed time-span, this can be removed for that period of (virtual) time. 
Ultimately the simulation could switch at runtime between protocols based on the information provided by activity tracking. This requires the above mentioned activity tracking framework, and could only be done if a GVT/LBTS time is agreed upon between the kernels. From the model point of view no changes are needed, although a non-trivial lookahead is obviously desired.
